\bf Swinging, Fast and Slow: Interpreting Variation in Baseball Swing Tracking Metrics

### 2025-12-30 articles/tas/paper.tex
move new figures from arxiv folder to output folder
clean up citations (#13)

\usepackage{amsfonts, tikz}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric}
\tikzset{>={To[length=3mm,width=3mm]}}
\graphicspath {{../../output/figures/}}
\newcommand{\tablepath}{../../output/tables/}
\newcommand{\inputtable}[1]{\input{\tablepath #1}}
\newcommand{\revision}[1]{{\color{orange} #1}}

@@

    Our dataset comprises 685,143 pitches thrown during the 2024 MLB regular season. This publicly available dataset was downloaded from Baseball Savant \citep{mlb_advanced_media_statcast_2024} using the \texttt{R} package \texttt{sabRmetrics} \citep{powers_sabrmetrics_2025}. For each pitch, we observe the identities and handedness (left/right) of the batter and pitcher involved and the ball-strike count prior to the pitch. We also observe pitch tracking information, most notably the pitch type (e.g. fastball, curveball, changeup) and the $(x, z)$ location where the pitch crosses the front of home plate. The pitch tracking information is sufficient to recreate an estimate of the pitch's full trajectory as a three-dimensional quadratic function of time $(x(t), y(t), z(t))$ and to recreate the corresponding pitch metrics such as release point, release speed, and vertical/horizontal break. Each pitch has a categorical outcome, which can be one of: Hit by Pitch (HBP), Called Ball, Called Strike, Swinging Strike, Foul Ball, or Fair Ball.

@@

      \revision{
        Table \ref{tab:intention-r2} summarizes the fraction of variance explained by the intention models based on the $R^2$ definition in \cite{gelman2019r}. Our model explains a higher fraction of variance in swing length compared to bat speed, and observe that most of the explained variation comes from the considered random effects for for both response variables.}

@@

      \begin{figure}[H]
        \centering
        \includegraphics[width = \textwidth]{articles/arxiv/figures/new_intent_re.pdf}
        \caption{\it Joint distribution of the posterior means for the batter-level $\mu_i$ random effects in the SK intention models for both bat speed and swing length. Each batter is displayed \revision{by points except for a select set of batters displayed by name}, and only batters with at least 25 squared-up swings are displayed.}
        \label{fig:intent-re}
      \end{figure}

@@

      \begin{figure}[H]
        \centering
        \includegraphics[width = 0.8\textwidth]{articles/arxiv/figures/new_alpha_re.pdf}
        \caption{\it Joint distribution of the posterior means for the batter-level $\alpha_i$ random effects in the SK intention models for both bat speed and swing length. Each batter is displayed \revision{by points except for a select set of batters displayed by name}, and only batters with at least 25 squared-up swings are displayed.}
        \label{fig:alpha-re}
      \end{figure}

@@

      \begin{figure}
        \centering
        \includegraphics[width = 0.8\textwidth]{articles/arxiv/figures/new_residual_sd.pdf}
        \caption{\it Joint distribution of the standard deviations of the swing length (y-axis) and bat speed (x-axis) residuals for all swings (not just squared up on primary fastballs) relative to the intention model predictions. Each batter is displayed by \revision{points except for a select set of batters displayed by name}, and only batters with at least 25 squared-up swings are displayed.}
        \label{fig:resid-sd}
      \end{figure}

>>>>

\usepackage{amsfonts, tikz}
\usetikzlibrary{arrows.meta, decorations.pathmorphing, positioning, shapes.geometric}
\tikzset{>={To[length=3mm,width=3mm]}}
\graphicspath {{../../output/figures/}}
\newcommand{\tablepath}{../../output/tables/}
\newcommand{\inputtable}[1]{\input{\tablepath #1}}
\newcommand{\revision}[1]{{\color{orange} #1}}

@@

    Our dataset comprises 685,143 pitches thrown during the 2024 MLB regular season. This publicly available dataset was downloaded from Baseball Savant \citep{baseball_savant_statcast_2024} using the \texttt{R} package \texttt{sabRmetrics} \citep{powers_sabrmetrics_2025}. For each pitch, we observe the identities and handedness (left/right) of the batter and pitcher involved and the ball-strike count prior to the pitch. We also observe pitch tracking information, most notably the pitch type (e.g. fastball, curveball, changeup) and the $(x, z)$ location where the pitch crosses the front of home plate. The pitch tracking information is sufficient to recreate an estimate of the pitch's full trajectory as a three-dimensional quadratic function of time $(x(t), y(t), z(t))$ and to recreate the corresponding pitch metrics such as release point, release speed, and vertical/horizontal break. Each pitch has a categorical outcome, which can be one of: Hit by Pitch (HBP), Called Ball, Called Strike, Swinging Strike, Foul Ball, or Fair Ball.

@@

      \revision{
        Table \ref{tab:intention-r2} summarizes the fraction of variance explained by the intention models based on the $R^2$ definition in \cite{gelman_r-squared_2019}. Our model explains a higher fraction of variance in swing length compared to bat speed, and observe that most of the explained variation comes from the considered random effects for for both response variables.}

@@

      \begin{figure}[H]
        \centering
        \includegraphics[width = \textwidth]{intent_re.pdf}
        \caption{\it Joint distribution of the posterior means for the batter-level $\mu_i$ random effects in the SK intention models for both bat speed and swing length. Each batter is displayed \revision{by points except for a select set of batters displayed by name}, and only batters with at least 25 squared-up swings are displayed.}
        \label{fig:intent-re}
      \end{figure}

@@

      \begin{figure}[H]
        \centering
        \includegraphics[width = 0.8\textwidth]{alpha_re.pdf}
        \caption{\it Joint distribution of the posterior means for the batter-level $\alpha_i$ random effects in the SK intention models for both bat speed and swing length. Each batter is displayed \revision{by points except for a select set of batters displayed by name}, and only batters with at least 25 squared-up swings are displayed.}
        \label{fig:alpha-re}
      \end{figure}

@@

      \begin{figure}
        \centering
        \includegraphics[width = 0.8\textwidth]{residual_sd.pdf}
        \caption{\it Joint distribution of the standard deviations of the swing length (y-axis) and bat speed (x-axis) residuals for all swings (not just squared up on primary fastballs) relative to the intention model predictions. Each batter is displayed by \revision{points except for a select set of batters displayed by name}, and only batters with at least 25 squared-up swings are displayed.}
        \label{fig:resid-sd}
      \end{figure}
